---
title: "Candy Hierarchy 2017"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE, warning = FALSE)
```

Thanks to David Ng (and Jenny Bryan for sharing) at UBC publishing the candy hierarchy data (http://www.scq.ubc.ca/so-much-candy-data-seriously/). One candy really stood out to me that I'm interested to learn more about it. Candy Corn, as the signature candy of Halloween, I have spent my life time questing the reason that people like it because I thought most people love it.

I'm glad to learn that I'm not alone on this. ___have mentioned it.

## Loading Packages and Reading in Data

```{r load, message=FALSE, error=FALSE}
library(tidyverse)
library(stringr)
library(jpeg)
library(grid)
# devtools::install_github('jbryer/likert')
library(likert)
library(reshape2)

img <- readJPEG("1468620.jpg") #Candy Corn background
data_orig <- read_csv("candyhierarchy2017.csv")
data      <- data_orig
colnames(data) <- colnames(data) %>%
  iconv(to = "ASCII//TRANSLIT") %>% #convert for special characters
  str_replace_all(c("\\\\xd5" = "",
                    "Q6 \\| |Q\\d{1,2}\\: " = "",
                    "\\?\\?" = ""))
# Re-order levels of Candy Corn for future plotting purposes
data$`Candy Corn` <- factor(data$`Candy Corn`, levels = c("DESPAIR", "MEH","JOY", NA))
```

## Feelings when Receive Candy Corn
First, more people feel despair about candy corn than joy.
NA means they left the question blank which can mean "they don't know the candy" as instructured in the survey or they skipped this option/question.

```{r pressure, echo=FALSE}
g <- ggplot(data, aes(`Candy Corn`))
g + annotation_custom(rasterGrob(img, width=unit(1,"npc"), height=unit(1,"npc")), 
                    -Inf, Inf, -Inf, Inf) +
  geom_bar(na.rm = TRUE, fill = "#ffdb99") +
  labs(title = "Feelings when Receive Candy Corn", x = "Feelings", y = "Frequency") +
  theme_minimal()
```

Now, let's do at these in percentages.
```{r}
cc_pert <- table(data$`Candy Corn`, useNA = "ifany") %>%
  prop.table() %>% #ignoring NAs
  data.frame() #%>%
  # filter(Freq > 0) # remove NA row
stderror <- function(p) {qnorm(.975) * sqrt(p * (1-p) / length(data$`Candy Corn`))} #95% CI
cc_pert$se <- sapply(cc_pert$Freq, stderror)

g <- ggplot(cc_pert, aes(x = Var1, y = Freq))
g + annotation_custom(rasterGrob(img, width=unit(1,"npc"), height=unit(1,"npc")), 
                      -Inf, Inf, -Inf, Inf) +
  geom_col(fill = "#ffdb99") +
  geom_errorbar(data=cc_pert, aes(ymin = Freq-se, ymax = Freq+se), width = 0.15, color = "dark grey") +
  geom_text(label = round(cc_pert$Freq, 3) * 100) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Feelings when Receive Candy Corn", x = "Feelings",
       subtitle = "With 95% Confidence Interval") +
  theme_minimal()
```

30.2% feel Despair if they receive candy corn, while only 19.4% feel joy. Surprisingly, 27.6% did not respond to the question, I am curious to know if they truly do not know what Candy Corn is, or if they decided to skip the question. Unfortunately, there is no way I can differentiate the two from the data.

## Likert Plot
Traditionaly, these data are usually plotted on a "Net Feelies" metric (=#JOY-#DESPAIR). Using likert plot can allow us to also take a look at the neutral option (MEH). 

```{r, fig.width = 15, fig.height=10}
candy <- data[, substr(colnames(data_orig), 1, 2) == "Q6"] #use index from orig data
candy2 <- candy %>%
  data.frame() %>%
  unclass() %>%
  as.data.frame()
lvls <- c("DESPAIR", "MEH","JOY", NA)
candy2[] <-  lapply(candy2, factor, levels=lvls)
colnames(candy2) <- colnames(data)[grep("Q6", colnames(data_orig))]

likert_plot <- likert(candy2)

p <- plot(likert_plot, colors = c("black", "grey", "orange")) + # change to Halloween color
  ggtitle(paste0("Candy Preference of N = ", nrow(candy2)))
print(p)
```

Wow, there are a lot of candies on this plot. In the likert package, there is no default option of shortening the plot by y-axis, and it will require some re-engineering of the function in order to shorten the plot.

# Candies that Receive >= 50% Joy
```{r highest joy}
# Re-engineer likert plot
# create summary table
table_summary <- likert(candy2)

# reshape results
results_50 <- table_summary$results %>% 
  filter(JOY >= 50) # Filter to 50% JOY
results <- melt(results_50, id.vars='Item')
results$Item <- factor(results_50$Item, levels = results_50$Item[order(results_50$JOY)]) #Sort bars from highest Joy

lsum <- summary(table_summary, center = (table_summary$nlevels-1)/2 + 1) %>%
  filter(high >= 50)
lsum$y <- lsum$low + (lsum$neutral/2)
# some defaults
ymin = 0
text.size = 3

ggplot(results, aes(y=value, x=Item, group=Item)) + 
  geom_bar(stat='identity', aes(fill=variable)) + 
  ylim(c(-5,105)) + 
  coord_flip() +
  scale_fill_manual('Response', values=c("black", "grey", "orange"), 
              breaks=levels(results$variable), 
              labels=levels(results$variable)) +
  geom_text(data=lsum, y=ymin, aes(x=Item, 
              label=paste(round(low), '%', sep='')), 
              size=text.size, hjust=1) +
  geom_text(data=lsum, y=100, aes(x=Item,
              label=paste(round(high), '%', sep='')), 
              size=text.size, hjust=-.2) +
  geom_text(data=lsum, aes(x=Item, y=y,
							  	   label=paste0(round(neutral), '%')),
							       size=text.size, hjust=.5) +
  labs(x = "", y = "Percentage", title = "Candy With >= 50% Joy")
```

# Candies Receiving Least % (<= 10%) of Joy
```{r least joy}
# reshape results
results_10 <- table_summary$results %>% 
  filter(JOY <= 10) # Filter to 50% JOY
results <- melt(results_10, id.vars='Item')
results$Item <- factor(results_10$Item, levels = results_10$Item[order(results_10$JOY)]) #Sort bars from highest Joy

lsum <- summary(table_summary, center = (table_summary$nlevels-1)/2 + 1) %>%
  filter(high <= 10)
lsum$y <- lsum$low + (lsum$neutral/2)
# some defaults
ymin = 0
text.size = 3

ggplot(results, aes(y=value, x=Item, group=Item)) + 
  geom_bar(stat='identity', aes(fill=variable)) + 
  ylim(c(-5,105)) + 
  coord_flip() +
  scale_fill_manual('Response', values=c("black", "grey", "orange"), 
              breaks=levels(results$variable), 
              labels=levels(results$variable)) +
  geom_text(data=lsum, y=ymin, aes(x=Item, 
              label=paste(round(low), '%', sep='')), 
              size=text.size, hjust=1) +
  geom_text(data=lsum, y=100, aes(x=Item,
              label=paste(round(high), '%', sep='')), 
              size=text.size, hjust=-.2) +
  geom_text(data=lsum, aes(x=Item, y=y,
							  	   label=paste0(round(neutral), '%')),
							       size=text.size, hjust=.5) +
  labs(x = "", y = "Percentage", title = "Candy With <= 10% Joy")
```

We can use the following interactive plot to zoom in or out on the plot.

The likert package also provides a look at the missing data for each candy item.
```{r missing data, fig.width = 15, fig.height=10}
plot(likert_plot, colors = c("black", "grey", "orange"), include.histogram = TRUE)
```

## Likert Plot by Age
It could also be interesting to learn the change of preference of candy with age.
First, we will clean up the age column by binning them.
```{r likert age}
# First, clean up age
data$age2 <- as.numeric(data$AGE)
# table(data$age2) #age 1, 312, 1000 and 39.4
data$age2[which(data$age2 >= 100 | data$age2 <= 1)] <- NA
# Second, bin age
agebreaks <- c(4, 9, 13, 18, 25, 35, 45, Inf)
agelabels <- c("4-8","9-12", "13-17", "18-24", "25-34", "35-44", "45+")
data$agebin <- cut(data$age2, breaks = agebreaks, right = FALSE, labels = agelabels)
```

Now, we are ready to plot!
```{r likert plot age, fig.width = 15, fig.height=10}
# Likert plot by age
plot_age <- function(candyage, agebin){
  candyage <- candyage %>%
    select(c(7:109)) %>%
    data.frame() %>%
    unclass() %>%
    as.data.frame()
  lvls <- c("DESPAIR", "MEH","JOY", NA)
  candyage[] <-  lapply(candyage, factor, levels=lvls)
  colnames(candyage) <- colnames(data)[grep("Q6", colnames(data_orig))]
  
  likert_plot_age4 <- likert(candyage)
  
  # reshape results
  table_summary <- likert_plot_age4
  results_50 <- table_summary$results %>% 
    filter(JOY >= 50) # Filter to 50% JOY
  results <- melt(results_50, id.vars='Item')
  results$Item <- factor(results_50$Item, levels = results_50$Item[order(results_50$JOY)]) #Sort bars from highest Joy
  
  lsum <- summary(table_summary, center = (table_summary$nlevels-1)/2 + 1) %>%
    filter(high >= 50)
  lsum$y <- lsum$low + (lsum$neutral/2)
  # some defaults
  ymin = 0
  text.size = 3
  
  p <- ggplot(results, aes(y=value, x=Item, group=Item)) + 
    geom_bar(stat='identity', aes(fill=variable)) + 
    ylim(c(-5,105)) + 
    coord_flip() +
    scale_fill_manual('Response', values=c("black", "grey", "orange"), 
                      breaks=levels(results$variable), 
                      labels=levels(results$variable)) +
    geom_text(data=lsum, y=ymin, aes(x=Item, 
                                     label=paste(round(low), '%', sep='')), 
              size=text.size, hjust=1) +
    geom_text(data=lsum, y=100, aes(x=Item,
                                    label=paste(round(high), '%', sep='')), 
              size=text.size, hjust=-.2) +
    geom_text(data=lsum, aes(x=Item, y=y,
                             label=paste0(round(neutral), '%')),
              size=text.size, hjust=.5) +
    labs(x = "", y = "Percentage", 
         title = paste0("Candy With >= 50% Joy of Age ", agebin, " (N = ", nrow(candyage), ")"))
  print(p)
  # print(plot(likert_plot_age4, colors = c("black", "grey", "orange")) + 
  #         ggtitle(paste0("Candy Preference of Age ", agebin)))  
}
# sapply(levels(data$agebin), plot_age) %>% invisible()
```

```{r, echo = FALSE}
candy_age4 <- data %>%
    filter(agebin == "4-8")
plot_age(candy_age4, "4-8")

candy_age9 <- data %>%
    filter(agebin == "9-12")
plot_age(candy_age9, "9-12")

candy_age13<- data %>%
    filter(agebin == "13-17")
plot_age(candy_age13, "13-17")

candy_age18 <- data %>%
    filter(agebin == "18-24")
plot_age(candy_age18, "18-24")

candy_age25 <- data %>%
    filter(agebin == "25-34")
plot_age(candy_age25, "25-34")

candy_age35 <- data %>%
    filter(agebin == "35-44")
plot_age(candy_age35, "35-44")

candy_age45 <- data %>%
    filter(agebin == "45+")
plot_age(candy_age45, "45+")
```

Here are some more plots I have for the purpose of exploratory data analysis.

# Should I give Candy Corn out on Trick or Treat? 
Up to you, it does not really matter to those who are going trick or treating.

```{r}
# Plot: Candy Corn and Going Out
ggplot(data, aes(x = factor(`GOING OUT?`), fill = factor(`Candy Corn`))) +
  geom_bar(na.rm = TRUE, position = position_dodge()) + 
  labs(title = "Feelings when Receive Candy Corn", x = "Going Trick or Treating?") +
  guides(fill=guide_legend(title="Feelings")) +
  theme_minimal()
```

```{r}
# Plot: Candy Corn and Age
ggplot(data, aes(x = factor(agebin), fill = factor(`Candy Corn`))) +
  geom_bar(na.rm = TRUE, position = position_dodge()) + 
  labs(title = "Feelings when Receive Candy Corn", x = "Age") +
  guides(fill=guide_legend(title="Feelings")) +
  theme_minimal()
```
